Interpretable Machine Learning

### 2025-05-17 papers/interpretable-machine-learning.qmd
Initial commit.

---
title: Papers
---

# Papers

## Recent Papers by Area

- [Deep Learning and Machine Learning](papers/deep-learning-and-machine-learning.qmd)
- [Veridical Data Science (PCS)](papers/veridical-data-science-pcs.qmd)
- [Interpretable Machine Learning](papers/interpretable-machine-learning.qmd)
- [Interdisciplinary Research in Biomedicine](papers/interdisciplinary-research-in-biomedicine.qmd)
- [Tree-based Methods](papers/tree-based-methods.qmd)

---

## Complete List by Reverse Chronological Order

*For the most up-to-date and complete list, see [Google Scholar](https://scholar.google.com/citations?user=xT19Jc0AAAAJ&hl=en) or [arXiv](https://arxiv.org/search/?query=bin+yu&searchtype=author).*

### Publications

1. A. R. Hsu, Y. Cherapanamjeri, A. Y. Odisho, P. R. Carroll, B. Yu (2024). [Mechanistic Interpretation through Contextual Decomposition in Transformers.](https://arxiv.org/pdf/2407.00886)
1. Y. S. Tan, O. Ronen, T. Saarinen, B. Yu (2024). [The Computational Curse of Big Data for Bayesian Additive Regression Trees: a Hitting Time Analysis.](https://arxiv.org/pdf/2406.19958)
1. O. Ronen, A. I. Humayun, R. Balestriero, R. Baraniuk, B. Yu (2024). [ScaLES: Scalable Latent Exploration Score for Pre-Trained Generative Networks](https://arxiv.org/pdf/2406.09657)
1. S. Hayou, N. Ghosh, B. Yu (2024). [The Impact of Initialization on LoRA Fineuning Dynamics.](https://arxiv.org/pdf/2406.08447)
1. B. Yu (2024). [After Computational Reproducibility: Scientific Reproducibility and Trustworthy AI](https://hdsr.mitpress.mit.edu/pub/8qexde24/release/1) (Harvard Data Science Review)
1. S. Hayou, N. Ghosh, B. Yu (2024). [LoRA+: Efficient Low Rank Adaptation of Large Models](https://arxiv.org/abs/2402.12354) (Proc. ICML)
1. C. F. Elliott, J. Duncan, T. M. Tang, M. Behr, K. Kumbier, B. Yu (2024). [Designing a data science simulation with MERITS: a primer.](https://arxiv.org/abs/2403.08971)
1. Y. Chen, C. Singh, X. Liu, S. Zuo, B. Yu, H. He, J. Gao (2024). [Towards consistent natural-language explanations via explanation-consistent finetuning.](https://arxiv.org/abs/2401.13986)
1. N. R. Mallinar, A. Zane, S. Frei, B. Yu (2024). Minimum-Norm Interpolation Under Covariate Shift. (Proc. ICML) [arXiv](https://arxiv.org/pdf/2404.00522)
1. L. Sun, A. Agarwal, A. Kornblith, B. Yu, C. Xiong (2024). ED-Copilot: Reduce Emergency Department Wait Time with Language Model Diagnostic Assistance. (Proc. ICML) [arXiv](https://arxiv.org/abs/2402.13448)
1. C. Singh*, A. R. Hsu*, R. Antonello, S. Jain, A. G. Huth, B. Yu, J. Gao (2023). [Explaining black box text modules in natural language with language models.](https://arxiv.org/abs/2305.09863)
1. Q. Zhang, C. Singh, L. Liu, X. Liu, B. Yu, J. Gao, T. Zhao (2023). [Tell your model where to attend: post-hoc attention steering for LLMs.](https://arxiv.org/abs/2311.02262)
1. Q. Wang*, T. M. Tang*, N. Youlton, C. S. Weldy, A. M. Kenney, O. Ronen, J. W. Hughes, E. T. Chin, S. C. Sutton, A. Agarwal, X. Li, M. Behr, K. Kumbier, C. S. Moravec, W. H. W. Tang, K. B. Margulies, T. P. Cappola, A. J. Buitte, R. Arnaout, J. B. Brown, J. R. Priest, V. N. Parikh, B. Yu*, E. Ashley* (2023). Epistasis regulates genetic control of cardiac hypertrophy. [medRxiv](https://www.medrxiv.org/content/10.1101/2023.11.06.23297858v1) ([Code](https://github.com/Yu-Group/epistasis-cardiac-hypertrophy)) ([PCS documentation](https://yu-group.github.io/epistasis-cardiac-hypertrophy/#overview))
1. R. Cahill, Y. Wang, R. P. Xian, A. J. Lee, H. Zeng, B. Yu, B. Tasic, R. Abbasi-Asl (2023). Unsupervised pattern discovery in spatial gene expression atlas reveals mouse brain regions beyond established ontology. [bioRxiv](https://www.biorxiv.org/content/10.1101/2023.03.10.531984v2) ([Code](https://github.com/abbasilab/osNMF))
1. E. Irajizad, A. Kenney, T. Tang, J. Vykoukal, R. Wu, E. Murage, J. B. Dennison, M. Sans, J. P. Long, M. Loftus, J. A. Chabot, M. D. Kluger, F. Kastrinos, L. Brais, A. Babic, K. Jajoo, L. S. Lee, T. E. Clancy, K. Ng, A. Bullock, J. M. Genkinger, A. Maitra, K. A. Do, B. Yu, B. W. Wolpin, S. Hanash, J. F. Fahrmann. (2023). [A blood-based metabolomic signature predictive of risk for pancreatic cancer](https://pubmed.ncbi.nlm.nih.gov/37729870/). Cell Reports Medicine 4(9): 101194. doi: 10.1016/j.xcrm.2023.101194. (PCS related) ([Editorial](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10518497/))
1. R. Dwivedi, C. Singh, B. Yu, M. Wainwright (2023). [Revisiting miniumu description length complexity in overparametrized models.](https://jmlr.org/papers/v24/21-1133.html) JMLR, 24(268): 1-59.
1. K. Wu, Y. Chen, W. Ha, B. Yu (2023). [Prominent roles of conditionally invariant components in domain adaptation: theory and algorithms.](https://arxiv.org/abs/2309.10301) JMLR (accepted). [https://arxiv.org/abs/2309.10301](https://arxiv.org/abs/2309.10301)
1. N. Ghosh, S. Frei, W. Ha, B. Yu (2023). The effect of SGD batch size on autoencoder learning: sparsity, sharpness and feature learning. [https://arxiv.org/abs/2308.03215](https://arxiv.org/abs/2308.03215)
1. R. Netzorg, J. Li, B. Yu (2024). Improving prototypical part networks with reward reweighting, reselection, and retraining. Proc. ICML. [https://arxiv.org/abs/2307.03887](https://arxiv.org/abs/2307.03887)
1. A. Agarwal, A. M. Kenny, Y. S. Tan, T. M. Tang, B. Yu (2023). MDI+: a flexible random forest-based feature importance framework. [https://arxiv.org/abs/2307.01932](https://arxiv.org/abs/2307.01932) (PCS related)
1. A. R. Hsu, Y. Cherapanamjeri, B. Park, T. Naumann, A. Odisho, and B. Yu (2023). Diagnosing transformers: illuminating feature space for clinical decison-making. Proc. ICLR. [https://arxiv.org/abs/2305.17588](https://arxiv.org/abs/2305.17588)
1. C. Singh, A. R. Hsu, R. Antonello, S. Jain, A. G. Huth, B. Yu and J. Gao (2023). Explaining black box text modules in natural language with language models. [https://arxiv.org/abs/2305.09863](https://arxiv.org/abs/2305.09863)
1. D. Shen, P. Ding, J. Sekhon, B. Yu (2022). Same root different leaves: time series and cross-sectional methods in panel data. Econometrica (accepted). [https://arxiv.org/abs/2207.14481](https://arxiv.org/abs/2207.14481)
1. B. Park, X. Wu, B. Yu, A. Zhou (2022). [Offline evaluation in RL: soft stability weighting to combine fitted Q-learning and model-based methods.](https://openreview.net/forum?id=R3h3-bt0mq) NeurIPS 3rd Offline Reinforcement Learning Workshop.
1. Y. S. Tan, C. Singh, K. Nasseri, A. Agarwal, J. Duncan, O. Ronen, M. Epland, A. Kornblith, B. Yu (2022). [Fast interpretable greedy-tree sums (FIGS)](https://arxiv.org/abs/2201.11931). ([imodels 🔎](https://github.com/csinva/imodels) contains code for FIGS)
1. A. Agarwal, Y. S. Tan, O. Ronen, C. Singh, B. Yu (2022). [Hierarchical shrinkage: improving accuracy and interpretability of tree-based methods.](https://arxiv.org/abs/2202.00858) Proc. ICML ([imodels 🔎](https://github.com/csinva/imodels) contains code for hierarchical shrinkage (HS)).
1. B. Yu and C. Singh (2022). [Seven principles for rapid-response data science: lessons learned from covid-19 forecasting](https://projecteuclid.org/journals/statistical-science/volume-37/issue-2/Seven-Principles-for-Rapid-Response-Data-Science--Lessons-Learned/10.1214/22-STS855.full). Statistical Science, 36(2):266-269.
1. N. Ghosh, S. Mei, and B. Yu (2022). [The three stages of dynamics in high-dimensional kernel methods.](https://arxiv.org/abs/2111.07167) Proc. ICLR, 2022.
1. Y. Tan, A. Agarwal, and B. Yu (2021). [A cautionary tale on fitting decision trees to data from additive models: generalization lower bounds.](https://arxiv.org/abs/2110.09626) Proc. AISTATS.
1. N. Altieri, B. Park, J. DeNero, A. Odisho, B. Yu. (2021). [Improving natural language information extraction from cancer pathology reports using transfer learning and zero-shot string similarity.](https://pubmed.ncbi.nlm.nih.gov/34604711/) JAMIA Open. 2021 Sept. 30 4(3).
1. C. Singh, W. Ha and B. Yu (2021). [Interpreting and Improving Deep-Learning Models with Reality Checks](https://arxiv.org/abs/2108.06847). To appear in "xxAI - Beyond Explainable AI" (eds. Holzinger et al.).
1. W. Ha, C. Singh, F. Lanusse, S. Upadhyayula, and B. Yu (2021). [Adaptive Wavelet Distillation from Neural Networks through Interpretation.](https://nips.cc/Conferences/2021/ScheduleMultitrack?event=28027) Proc. NeurIPS 2021. ([code](https://github.com/Yu-Group/adaptive-wavelets))
1. M. Behr, Y. Wang, X. Li, B. Yu (2022). [Provable Boolean Interaction Recovery from Tree Ensemble obtained via Random Forests.](https://arxiv.org/abs/2102.11800) PNAS. ([theory for iRF](https://www.pnas.org/doi/10.1073/pnas.1711236115), PCS-related)
1. N. Altieri, B. Park, M. Olson, J. DeNero, A. Odisho, B. Yu. (2021). Supervised line attention for tumor attribute classification from pathology reports: Higher performance with less data. Journal of Biomedical Informatics. 122 (2021) 103872. [arXiv version](https://arxiv.org/abs/2012.08113)
1. M. Behr*, K. Kumbier*, A. Cordova-Palomera, M. Aguirre, E. Ashley, A. Butte, R. Arnaout, J. B. Brown, J. Preist*, B. Yu* (2020). [Learning epistatic polygenic phenotypes with Boolean interactions](https://www.biorxiv.org/content/10.1101/2020.11.24.396846v1) ([code](https://github.com/merlebehr/epiTree)) (PCS inference case study)
1. B. Norgeot*, G. Quer, B. K. Beaulieu-Jones, A. Torkamani, R. Dias, M. Gianfrancesco, R. Arnaout, I. S. Kohane, S. Saria, E. Topol, Z. Obermeyer, B. Yu & A. Butte* (2020). [Minimum information about clinical artificial intelligence modeling: the MI-CLAIM checklist](https://www.nature.com/articles/s41591-020-1041-y), Nature Medicine, 26, 1320–1324.
1. B. Yu (2020). [Stability expanded, in reality](https://hdsr.mitpress.mit.edu/pub/ekrhsui8/release/1). Harvard Data Science Review (PCS related)
1. B. Yu and R. Barter (2020). [Data science process: one culture.](https://www.tandfonline.com/doi/abs/10.1080/01621459.2020.1762615?journalCode=uasa20) JASA. (PCS related)
1. R. Dwivedi, Y. Tan, B. Park, M. Wei, K. Horgan, D. Madigan, B. Yu (2020). [Stable discovery of interpretable subgroups via calibration in causal studies (staDISC).](https://onlinelibrary.wiley.com/doi/10.1111/insr.12427) International Statistical Review ([code](https://github.com/Yu-Group/stadisc)) (PCS case study for causal inference)
1. X. Li, T. M. Tang, X. Wang, J. A. Kocher, B. Yu (2020). [A stability-driven protocol for drug response interpretable prediction (staDRIP)](https://arxiv.org/abs/2011.06593). NeurISP workshop on ML4H (Machine learning for Health) Extended Abstract.
1. A. Y. Odisho, B. Park, N. Altieri, J. DeNero, M. R Cooperberg, P. R .Carroll, B. Yu (2020). [Natural language processing systems for pathology parsing in limited data environments with uncertainty estimation.](https://academic.oup.com/jamiaopen/article/3/3/431/5922788) JAMIA Open.
1. L. Reiger, J. W. Murdoch, S. Singh, B. Yu (2020). [Interpretations are Useful: Penalizing Explanations to Align Neural Networks with Prior Knowledge](https://proceedings.icml.cc/static/paper_files/icml/2020/992-Paper.pdf). ICML Proceedings. ([code](https://github.com/laura-rieger/deep-explanation-penalization))
1. C. Singh, W. Ha, F. Lanusse, V. Boehm, J. Liu, B. Yu (2020). [Transformation Importance with Applications to Cosmology](https://arxiv.org/pdf/2003.01926.pdf) ICLR Workshop paper. ([code](https://github.com/csinva/transformation-importance))
1. N. Altieri, R. Barter, J. Duncan, R. Dwivedi, K. Kumbier, X. Li, R. Netzorg, B. Park, C. Singh*, Y. Tan, T.Tang, Y. Wang, C. Zhang, B. Yu*. (2020) [Curating a COVID-19 data repository and forecasting county-level death counts in the United States.](https://hdsr.mitpress.mit.edu/pub/p6isyf0g/release/1) Harvard Data Science Review ([code](https://github.com/Yu-Group/covid19-severity-prediction)) [7-day prediction results](https://covidseverity.com/) [Short talk video](https://www.stat.berkeley.edu/~binyu/ps/papers2020/RDSS20-yu.mp4)
1. B. Yu and K. Kumbier (2020) [Veridical data science](https://www.stat.berkeley.edu/~binyu/ps/papers2020/VDS20-YuKumbier.pdf) (PCS framework), PNAS. 117 (8), 3920-3929. [QnAs with Bin Yu.](https://www.stat.berkeley.edu/~binyu/ps/papers2020/VDS20-QnAsBinYu.pdf)
1. R. Dwivedi, N. Ho, K. Khamaru, M. J. Wainwright, M. I. Jordan and B. Yu (2020) [Sharp Analysis of Expectation-Maximization for Weakly Identifiable Mixture Models](https://arxiv.org/abs/1902.00194) AISTATS.
1. R. Dwivedi, N. Ho, K. Khamaru, M. J. Wainwright, M. I. Jordan and B. Yu (2020) [Singularity, Misspecification and the Convergence Rate of EM](https://arxiv.org/abs/1810.00828) Annals of Statistics.
1. Y. Chen, R. Dwivedi, M. J. Wainwright and B. Yu (2020) [Fast Mixing of Metropolized Hamiltonian Monte Carlo: Benefits of Multi-Step Gradients,](https://www.jmlr.org/papers/volume21/19-441/19-441.pdf) JMLR, [arXiv](https://arxiv.org/abs/1905.12247)
1. R. Dwivedi, Y. Chen, M. J. Wainwright and B. Yu (2019) [Log-concave Sampling: Metropolis Hastings Algorithms are Fast](http://jmlr.org/papers/v20/19-306.html) JMLR.
1. D. Rothenhäusler and B. Yu (2019). [Incremental causal effects](https://arxiv.org/abs/1907.13258).
1. Y. Chen, R. Dwivedi, M. J. Wainwright and B. Yu (2018) [Fast MCMC Algorithms on Polytopes](http://jmlr.org/papers/v19/18-158.html). JMLR . http://jmlr.org/papers/v19/18-158.html
1. Y. Chen, R. Dwivedi, M. J. Wainwright and B. Yu (2020) [Vaidya Walk: A Sampling Algorithm Based on Volumetric-Logarithmic Barrier](https://ieeexplore.ieee.org/abstract/document/8262876/) Allerton Conference 2017 https://ieeexplore.ieee.org/abstract/document/8262876/
1. W. J. Murdoch, C. Singh, K. Kumbier, R. Abbasi-Asl, and B. Yu* (2019) [Definitions, methods, and applications in interpretable machine learning.](https://www.stat.berkeley.edu/~binyu/ps/papers2020/iML19-Murdochetal.pdf) PNAS, 116 (44) 22071-22080.
1. W. J. Murdoch, C. Sign, and B. Yu (2019). [Hierarchical interpretations for neural network predictions.](https://openreview.net/pdf?id=SkEqro0ctQ) ICLR. ([code](https://github.com/csinva/hierarchical-dnn-interpretations))
1. Y. Wang, S. Wu and B. Yu (2020) [Unique Sharp Local Minimum in l1-minimization Complete Dictionary Learning.](https://jmlr.org/papers/volume21/19-169/19-169.pdf) JMLR. 21(63), pp. 1-52. Also at https://arxiv.org/abs/1902.08380
1. Y. Chen, R. Abbasi-Asl, A. Bloniarz, M. Oliver, B. Willmore, J. Gallant*, and B. Yu* (2018) [The DeepTune framework for modeling and characterizing neurons in visual cortex area V4](https://www.biorxiv.org/content/10.1101/465534v1) https://www.biorxiv.org/content/10.1101/465534v1
1. K. Kumbier, S. Sumanta, J. B. Brown, S. Celniker, and B. Yu* (2018) [Refining interaction search through signed iterative Random Forests.](https://arxiv.org/abs/1810.07287) [https://arxiv.org/abs/1810.0728](https://arxiv.org/abs/1810.0728) (an enhanced version of [iRF](https://www.pnas.org/doi/10.1073/pnas.1711236115), PCS related)
1. Y. Chen C. Jin, and B. Yu (2018) [Stability and Convergence Trade-off of Iterative Optimization Algorithms.](https://arxiv.org/abs/1804.01619) https://arxiv.org/abs/1804.01619
1. J. Murdoch, P. Liu, and B. Yu (2018) [Beyond word importance: contextual decomposition to extract interactions from LSTMs.](https://arxiv.org/abs/1801.05453) Proc. ICLR 2018. [https://arxiv.org/abs/1705.07356](https://arxiv.org/abs/1705.07356) ([code)](https://github.com/csinva/hierarchical-dnn-interpretations)
1. R. Diwivedi, Y. Chen, M. J. Wainwright, and B. Yu (2018) [Log-concave sampling: Metropolis-Hastings algorithms are fast!](https://arxiv.org/abs/1801.02309) https://arxiv.org/abs/1801.02309.
1. Y. Chen, R. Dwivedi, M. J. Wainwright, and B. Yu (2017) [Fast MCMC sampling algorithms on polytopes](https://arxiv.org/abs/1710.08165) https://arxiv.org/abs/1710.08165.
1. B. Yu and K. Kumbier (2018) [Artificial Intelligence and Statistics](https://www.stat.berkeley.edu/~binyu/ps/papers2018/AI+Stat18.pdf) Frontiers of Information Technology and Electronic Engineering. 19(1), 6-9.
1. R. Abbasi-Asl and B. Yu (2017) [Structural Compression of Convolutional Neural Networks Based on Greedy Filter Pruning](https://arxiv.org/abs/1705.07356) https://arxiv.org/abs/1705.07356
1. R. Abbasi-Asl and B. Yu (2017) [Interpreting Convolutional Neural Networks Through Compression.](https://arxiv.org/abs/1711.02329) NIPS 2017. Symposium on Interpretable Machine Learning. (also https://arxiv.org/abs/1711.02329)
1. S. Kunzel, J. Sekhon, P. Bickel, and B. Yu* (2019) [Meta-learners for Estimating Heterogeneous Treatment Effects using Machine Learning,](https://www.pnas.org/content/116/10/4156) PNAS. 116 (10) 4156-4165.  https://arxiv.org/abs/1706.03461. [(code](https://github.com/soerenkuenzel/hte))
1. S. Basu, K. Kumbier, J. B. Brown*, and B. Yu* (2018) [iterative Random Forests to discover predictive and stable high-order interactions](https://www.stat.berkeley.edu/~binyu/ps/papers2018/iRF+SI18.pdf) PNAS, 115 (8), 1943-1948. ([code](https://github.com/Yu-Group/iterative-Random-Forest)) (PCS related)
1. S. Balakrishnan, M. Wainwright, B. Yu (2017) [Statistical Guarantees for the EM algorithm: from population to sample-based analysis.](https://www.stat.berkeley.edu/~binyu/ps/papers2017/EM17.pdf) Annals of Statistics, 45(1), 77 - 120.
1. R. Barter and B. Yu (2017) [Superheat: An R package for creating beautiful and extendable heatmaps for visualizing complex data](https://www.stat.berkeley.edu/~binyu/ps/papers2017/superheat17.pdf) JCGS (revised). https://github.com/rlbarter/superheat
1. H. Liu and B. Yu (2017) [Comments on: High-dimensional simultaneous inference with the bootstrap by Dezeure et al](https://www.stat.berkeley.edu/~binyu/ps/papers2017/LiuYu-test17.pdf) Test. 26: 740-750.
1. C. Carson et al (2016). [UC Berkeley Data Science Planning Initiative Faculty Advisory Board (FAB) Report.](https://www.stat.berkeley.edu/~binyu/ps/FAB2016.pdf) [FAB Report Executive Summary](https://www.stat.berkeley.edu/~binyu/ps/FAB-Summary2016.pdf)
1. S. Wu and B. Yu (2018). [Local identifiability of l1-minimization dictionary learning: a sufficient and almost necessary condition.](https://www.stat.berkeley.edu/~binyu/ps/papers2018/local-iden18.pdf) JMLR. 18, 1 - 56.
1. K. Rohe, T. Qin and B. Yu* (2016). [Co-clustering directed graphs to discover asymmetries and directional communities.](http://www.pnas.org/content/113/45/12679.full.pdf?with-ds=yes) Proc. National Academy of Sciences (PNAS), 113(45), 12679 - 12684.
1. R. E. Kass, B. S. Caffo, M. Davidian, X. Meng, B. Yu, Nancy Reid* (2016). [Ten simple rules for effective statistical practice.](https://www.stat.berkeley.edu/~binyu/ps/papers2016/Ten-Simple-Rules.pdf) PLoS Comput. Biol., 12(6): e1004961. doi:10.1371/journal.pcbi.1004961
1. Siqi Wu, Antony Joseph, Ann S. Hammonds, Susan E. Celniker, Bin Yu*, and Erwin Frise* (2016). [Stability-driven nonnegative matrix factorization to interpret spatial gene expression and build local gene networks (with support information).](https://www.stat.berkeley.edu/~binyu/ps/papers2016/staNMF+SI.pdf) PNAS, pp. 4290 - 4295. ([code](https://github.com/yu-group/stanmf)) (PCS related)
1. A. Bloniarz, C. Wu, B. Yu, A. Talwalkar (2016). [Supervised neighborhoods for distributed nonparametric regression.](https://www.stat.berkeley.edu/~binyu/ps/papers2016/AISTATS16.pdf) Proc. of AISTATS, Barcelona, Spain.
1. B. Yu (2015). [Data wisdom for data science.](http://www.odbms.org/2015/04/data-wisdom-for-data-science/) Operational Database Management Systems (ODBMS.ORG).
1. A. Bloniarz, H. Liu, C. Zhang, J. Sekhon, and B. Yu* (2015). [Lasso adjustments of treatment effect estimates in randomized experiments.](https://www.stat.berkeley.edu/~binyu/ps/papers2016/lasso-adj+SI16.pdf) PNAS. 113, 7383 - 7390.
1. P. Ma, M. W. Mahoney and B. Yu (2015). [A Statistical Perspective on Algorithmic Leveraging.](https://www.stat.berkeley.edu/~binyu/ps/papers2015/MaMahoneyYu2015.pdf) Journal of Machine Learning Research, 16, (2015), 861-911.
1. T. Moon, Y. Wang, Y. Liu, and B. Yu (2015). [Evaluation of a MISR-based high-resolution aerosol retrieval method using AERONET DRAGON campaign data.](http://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=7047927) IEEE Transactions on Geoscience and Remote Sensing, 53, 4328-4339.
1. B. Yu (2014). [Let us own data science.](https://www.stat.berkeley.edu/~binyu/ps/papers2014/IMS-pres-address14-yu.pdf) [video](https://www.youtube.com/watch?v=92OjsYQJC1U) [IMS Bulletin](http://bulletin.imstat.org/2014/10/ims-presidential-address-let-us-own-data-science/) Institute of Mathematical Statistics (IMS) Presidental Address, ASC-IMS Joint Conference, Sydney, July, 2014.
1. G. Schiebinger, M. J. Wainwright and B. Yu (2014). [The geometry of kernelized spectral clustering.](http://arxiv.org/pdf/1404.7552v3.pdf) Annals of Statistics, 43, 819-846.
1. L. Miratrix, J. Jia, B. Yu, B. Gawalt, L. El Ghaoui, L. Barnesmoore, S. Clavier (2014). [Concise comparative summaries (CCS) of large text corpora with a human experiment.](https://www.stat.berkeley.edu/~binyu/ps/papers2014/CCS14.pdf) Ann. Applied Statist., 8, 499-529.
1. Y. Benjamini and B. Yu (2014). [The shuttle estimator for explainable variance in fMRI experiments.](https://www.stat.berkeley.edu/~binyu/ps/papers2012/BenjaminiY12.pdf) Annals of Applied Statistics, 7, 2007-2033.
1. D. Bean, P. Bickel, N. El Karoui and B. Yu (2014). [Optimal M-estimation in high-dimensional regression.](http://www.pnas.org/content/early/2013/08/08/1307845110.full.pdf+html) Proceedings of National Academy of Sciences, 110, 1456314568.
1. N. El Karoui, D. Bean, P. Bickel, C. Lim, and B. Yu (2014). [On robust regression with high-dimensional predictors.](http://www.pnas.org/content/early/2013/08/15/1307842110.full.pdf+html) Proceedings of National Academy of Sciences, 110, 1455714562.
1. P. Ma, M. W. Mahoney, B. Yu (2014). [A Statistical Perspective on Algorithmic Leveraging.](http://arxiv.org/abs/1306.5362) Proc. of International Conference on Machine Learning (ICML) (This conference paper contains some of preliminary results of the journal submission Ma et al. (2015))
1. A. Bloniarz, A. Talwalkar, J. Terhorst, M. Jordan, D. Patterson, B. Yu and Y. Song (2014). [Changepoint Analysis for Efficient Variant Calling.](https://www.stat.berkeley.edu/~binyu/ps/papers2014/Bloniarzetal14.pdf) Proc. of RECOMB 2014 (to appear).
1. Tao Shi (2013), [A conversation with Professor Bin Yu](http://www.yufamily.org/src/ShiTao_AnInterviewWithProfessorBinYu.php?xref=YuFamily_YuBin_interview_cn) International Chinese Statistical Association (ICSA) Bulletin, Vol 25, Issue 2, pp 85-98. ([Selected Parts in Statblogs](http://www.statsblogs.com/2013/07/09/a-conversation-with-professor-bin-yu/))
1. A. Joseph and B. Yu (2016). [The impact of regularization on spectral clustering.](https://www.stat.berkeley.edu/~binyu/ps/papers2016/reg-spectral16.pdf) Annals of Statistics. 4, 1765 - 1791.
1. C. Lim and B. Yu (2016). [Estimation Stability with Cross Validation (ESCV)](https://www.stat.berkeley.edu/~binyu/ps/papers2016/ESCV16.pdf) Journal of Computational and Graphical Statistics. 25, 464 - 492. (First paper towards PCS)
1. A. S. Hammonds, C. A. Bristow, W. W. Fisher, R. Weiszmann, S. Wu, V. Hartenstein, M. Kellis, B. Yu, E. Frise, and S. E. Celniker (2013). [Spatial expression of transcription factors in Drosophila embryonic organ development.](https://www.stat.berkeley.edu/~binyu/ps/papers2013/Hammondsetal13.pdf) Genome Biology, 14(12), R140.
1. H. Liu and B. Yu (2013). [Asymptotic properties of Lasso+mLS and Lasso+Ridge in sparse high-dimensional linear regression.](https://www.stat.berkeley.edu/~binyu/ps/LassoOLSBootstrap13.pdf) Electron. J. Statist., 7, 312-3169.
1. J. Mairal and B. Yu (2013). [Supervised Feature Selection in Graphs with Path Coding Penalties and Network Flows.](http://jmlr.org/papers/v14/mairal13a.html) Journal of Machine Learning Research, 14, 2449-2485.
1. Y. Wang, X. Jiang, B. Yu, M. Jiang (2013). [A Hierarchical Bayesian Approach for Aerosol Retrieval Using MISR Data.](https://www.stat.berkeley.edu/~binyu/ps/papers2013/WangJiangYuJiang2013.pdf) J. American Statistical Association, 108, 483-493.
1. Y. He, J. Jia and B. Yu (2013). [Reversible MCMC on Markov equivalence classes of sparse directed acyclic graphs.](https://www.stat.berkeley.edu/~binyu/ps/papers2013/rev-MCMC13.pdf) Annals of Statistics, 41(4), 1742-1779.
1. B. Yu (2013). [Stability.](https://www.stat.berkeley.edu/~binyu/ps/papers2013/Yu13.pdf) Bernoulli, 19 (4), 1484-1500. (Invited paper for the Special Issue commemorating the 300th anniversary of the publication of Jakob Bernoullis Ars Conjectandi in 1712) (Begining of PCS)
1. J. Mairal and B. Yu (2013). Discussion on Grouping Strategies and Thresholding for High Dimensional Linear Models Journal of Statistical Planning and Inference, 143, 1451-1453.
1. C. Uhler, G. Raskutti, and P. Buhlmann and B. Yu (2013). [Geometry of faithfulness assumption in causal inference.](https://www.stat.berkeley.edu/~binyu/ps/papers2013/faithfulness13.pdf) Annals of Statistics, 41, 436-463.
1. L. Miratrix, J. Sehkon, and B. Yu (2013). [Adjusting Treatment Effect Estimates by Post-Stratification in Randomized Experiments.](https://www.stat.berkeley.edu/~binyu/ps/papers2013/post-stratification13.pdf) Journal of Royal Statistical Society, Series B, 75 (part 2), 369-396.
1. J. Jia, K. Rohe and B. Yu (2013) [The Lasso under Poisson-like Heteroscadecity.](https://www.stat.berkeley.edu/~binyu/ps/papers2013/lasso-poisson13.pdf) Statistica Sinica, 23, 99-118.
1. S. Negahban, P. Ravikumar, M. Wainwrigt, and B. Yu (2012) [A unified framework for high-dimensional analysis of M-estimators with decomposable regularizers.](https://www.stat.berkeley.edu/~binyu/ps/unified-797.pdf) Statistical Science, 27, 538-557.
1. G. Raskutti, M. Wainwrigt, and B. Yu (2012) [Minimax-optimal rates for sparse additive models over kernel classes via convex programming.](https://www.stat.berkeley.edu/~binyu/ps/papers2012/RaskuttiWY12.pdf) J. Machine Learning Research, 13, 389-427.
1. J. Mairal and B. Yu (2012). [Complexity analysis of the Lasso regularization path.](https://www.stat.berkeley.edu/~binyu/ps/Conferencepapers/MairalY12.pdf) Proc. of International Conference on Machine Learning (ICML).

>>>>

---
title: "Interpretable Machine Learning"
---

[Complete Paper List in Reverse Chronological Order](../papers.qmd)

## Selected Recent Papers on Interpretable Machine Learning

1. A. R. Hsu, Y. Cherapanamjeri, A. Y. Odisho, P. R. Carroll, B. Yu (2024). [Mechanistic Interpretation through Contexual Decomposition in Transformers.](https://arxiv.org/pdf/2407.00886) [https://arxiv.org/pdf/2407.00886](https://arxiv.org/pdf/2407.00886).
1. Y. Chen, C. Singh, X. Liu, S. Zuo, B. Yu, H. He, J. Gao (2024). [Towards consistent natural-language explanations via explanation-consistent finetuning.](https://arxiv.org/abs/2401.13986) [https://arxiv.org/abs/2401.13986](https://arxiv.org/abs/2401.13986)
1. Q. Zhang, C. Singh, L. Liu, X. Liu, B. Yu, J. Gao, T. Zhao (2023). [Tell your model where to attend: post-hoc attention steering for LLMs.](https://arxiv.org/abs/2311.02262) ICLR 2024. https://arxiv.org/abs/2311.02262
1. A. Agarwal, A. M. Kenny, Y. S. Tan, T. M. Tang, B. Yu (2023). MDI+: a flexible random forest-based feature importance framework. [https://arxiv.org/abs/2307.01932](https://arxiv.org/abs/2307.01932) (PCS related)
1. A. R. Hsu, Y. Cherapanamjeri, B. Park, T. Naumann, A. Odisho, and B. Yu (2023). [Diagnosing transformers: illuminating feature space for clinical decison-making.](https://arxiv.org/abs/2305.17588) ICLR (2024) [https://arxiv.org/abs/2305.17588](https://arxiv.org/abs/2305.17588)
1. C. Singh, A. R. Hsu, R. Antonello, S. Jain, A. G. Huth, B. Yu and J. Gao (2023). [Explaining black box text modules in natural language with language models.](https://arxiv.org/abs/2305.09863)
1. C. Singh, W. Ha and B. Yu (2021). [Interpreting and Improving Deep-Learning Models with Reality Checks](https://arxiv.org/abs/2108.06847). [https://arxiv.org/abs/2108.06847](https://arxiv.org/abs/2108.06847) to appear in the book entitled "xxAI - Beyond Explainable AI" (eds. Andreas Holzinger, Randy Goebel, Ruth Fong, Taesup Moon, Klaus-Robert Müller, and Wojciech Samek).
1. W. Ha, C. Singh, F. Lanusse, S. Upadhyayula, and B. Yu (2021). [Adaptive Wavelet Distillation from Neural Networks through Interpretation.](https://nips.cc/Conferences/2021/ScheduleMultitrack?event=28027) Proc. NeurIPS 2021. ([code](https://github.com/Yu-Group/adaptive-wavelets))
1. L. Reiger, J. W. Murdoch, S. Singh, B. Yu (2020). [Interpretations are Useful: Penalizing Explanations to Align Neural Networks with Prior Knowledge](https://proceedings.icml.cc/static/paper_files/icml/2020/992-Paper.pdf). ICML Proceedings. ([code](https://github.com/laura-rieger/deep-explanation-penalization))
1. C. Singh, W. Ha, F. Lanusse, V. Boehm , J. Liu, B. Yu (2020). [Transformation Importance with Applications to Cosmology](https://arxiv.org/pdf/2003.01926.pdf) ICLR Workshop paper. ([code](https://github.com/csinva/transformation-importance))
1. W. J. Murdoch, C. Singh, K. Kumbier, R. Abbasi-Asl, and B. Yu* (2019) [Definitions, methods, and applications in interpretable machine learning.](https://www.stat.berkeley.edu/~binyu/ps/papers2020/iML19-Murdochetal.pdf) PNAS, 116 (44) 22071-22080.
1. W. J. Murdoch, C. Sign, and B. Yu (2019). [Hierarchical interpretations for neural network predictions.](https://openreview.net/pdf?id=SkEqro0ctQ) ICLR. ([code](https://github.com/csinva/hierarchical-dnn-interpretations))
1. J. Murdoch, P. Liu, and B. Yu (2018) [Beyond word importance: contextual decomposition to extract interactions from LSTMs.](https://arxiv.org/abs/1801.05453) Proc. ICLR 2018. [https://arxiv.org/abs/1705.07356](https://arxiv.org/abs/1705.07356) ([code](https://github.com/csinva/hierarchical-dnn-interpretations))
